Boolean Operations for Surfaces in VTK Without External Libraries

### 2011-05-10 Document/LaTeX/Article.tex
Added results and discussion about how to use classes to compute surfaces from boolean operations.

\usepackage[dvips]{graphicx}

@@

While it seems possible to use the class \code{vtkClipPolyData} together with \code{vtkImplicitPolyData} to extract the necessary portions of each input surface to produce the boolean surfaces, there is a problem. \code{vtkClipPolyData} treats the implicit function as piecewise linear within each cell of the \code{vtkPolyData} being clipped. In general, implicit functions are not guaranteed to be piecewise linear in each cell, and the same is true for the signed distance field. The result is that clipped surfaces produced by \code{vtkClipPolyData} will not be clipped at the true zero level of the distance field. Consequently, the clipped portions of each input geometry will not match at the zero level where they should fuse seamlessly. XXXX Figure

@@

@@

XXXX Show pictures of various objects with boolean operations applied to them.

@@

The source code contributed with this article has been build against VTK 5.8.

>>>>

\usepackage[dvips]{graphicx}
\usepackage{subfigure}

@@

While it seems possible to use the class \code{vtkClipPolyData} together with \code{vtkImplicitPolyData} to extract the necessary portions of each input surface to produce the boolean surfaces, there is a problem. \code{vtkClipPolyData} treats the implicit function as piecewise linear within each cell of the \code{vtkPolyData} being clipped. In general, implicit functions are not guaranteed to be piecewise linear in each cell, and the same is true for the signed distance field. The result is that clipped surfaces produced by \code{vtkClipPolyData} will not be clipped at the true zero level of the distance field. Consequently, the clipped portions of each input geometry will not match at the zero level where they should fuse seamlessly.

\begin{figure}[t]
\centering
\includegraphics[scale=0.25]{Figures/DistanceZeroLevelVsIntersection}
\itkcaption{A case where the zero level of the distance field from the cylinder (black lines) does not match the intersection between the cone and the cylinder (white lines). Because of this mismatch, the class \code{vtkClipPolyData} is not suitable for accurate splitting of one surface mesh by another.}
\label{fig:DistanceZeroLevelVsIntersection}
\end{figure}

@@

\section{Putting it All Together}

To compute a surface resulting from a boolean operation on surface meshes, the following pipeline of classes can be used to compute the desired surface:

\begin{center}
\code{vtkPolyDataIntersection}

$\downarrow$

\code{vtkPolyDataDistance}

$\swarrow$ \hspace{3em} $\searrow$

\code{vtkThreshold} \code{vtkThreshold}

$\downarrow$ \hspace{5em}  $\downarrow$

\code{vtkDataSetSurfaceFilter} \code{vtkDataSetSurfaceFilter}

$\searrow$ \hspace{3em} $\swarrow$

\code{vtkAppendPolyData}
\end{center}

The two instances of \code{vtkThreshold} and two instances of \code{vtkDataSetSurfaceFilter} are required to select different portions from each of the input geometries.

For convenience, we have included a single class that replicates the functionality  class for computing boolean operations on surface meshes, \code{vtkPolyDataBooleanOperationFilter}, that encapsulates the functionality of the pipeline above.

@@

Results from boolean operations on various pairs of geometric objects are shown in Figure \ref{fig:Results}.

\input{ResultsFigure.tex}

@@

The source code contributed with this article has been built against VTK 5.8.
